---
title: "Conte Data Analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---
Step 1. Read in the data 
```{r}
library(readr)
data1 <- read_csv("/Volumes/klie/data/projects/conte/code/r_analysis/conte_data.csv")
```

Step 2. Perform z-transform/standardization on data to make variables comparable (adjusts mean and standard deviation). 

```{r}
library(robustHD)
data2 <- NULL
data2$age_mean_c <- standardize(x = data1$age, centerFun = mean, scaleFun = sd)
hist(data2$age_mean_c)
data2$sex_mean_c <- standardize(x = data1$sex, centerFun = mean, scaleFun = sd)
hist(data2$sex_mean_c)
data2$handedness_mean_c <- standardize(x = data1$handedness, centerFun = mean, scaleFun = sd)
hist(data2$handedness_mean_c)
data2$fsiq_mean_c <- standardize(x = data1$fsiq, centerFun = mean, scaleFun = sd)
hist(data2$fsiq_mean_c)
data2$eq_mean_c <- standardize(x = data1$eq, centerFun = mean, scaleFun = sd)
hist(data2$eq_mean_c)
data2$lins_etiv_mean_c <- standardize(x = data1$lins_etiv, centerFun = mean, scaleFun = sd)
hist(data2$lins_etiv_mean_c)
data2$rins_etiv_mean_c <- standardize(x = data1$rins_etiv, centerFun = mean, scaleFun = sd)
hist(data2$rins_etiv_mean_c)
data2$lifg_etiv_mean_c <- standardize(x = data1$lifg_etiv, centerFun = mean, scaleFun = sd)
hist(data2$lifg_etiv_mean_c)
data2$rifg_etiv_mean_c <- standardize(x = data1$rifg_etiv, centerFun = mean, scaleFun = sd)
hist(data2$rifg_etiv_mean_c)
data2$lracc_etiv_mean_c <- standardize(x = data1$lracc_etiv, centerFun = mean, scaleFun = sd)
hist(data2$lracc_etiv_mean_c)
data2$rracc_etiv_mean_c <- standardize(x = data1$rracc_etiv, centerFun = mean, scaleFun = sd)
hist(data2$rracc_etiv_mean_c)
data2$lvmpfc_etiv_mean_c <- standardize(x = data1$lvmpfc_etiv, centerFun = mean, scaleFun = sd)
hist(data2$lvmpfc_etiv_mean_c)
data2$rvmpfc_etiv_mean_c <- standardize(x = data1$rvmpfc_etiv, centerFun = mean, scaleFun = sd)
hist(data2$rvmpfc_etiv_mean_c)
```

Step 3. Run linear models 

Independent/Predictor Variable: Empathizing Quotient (x)
Dependent/Predicted Variable: Cortical Thickness (y)
Covariates: age, sex, handedness, and fsiq (fsiq included in data fit2 as an exploratory covariate). 
  Question: In talking with Marco Pipoly, he wondered if instead of using fsiq (a fluid IQ measure), if we should use verbal IQ (a crystalized IQ measure), based on literature findings. https://www.sciencedirect.com/topics/neuroscience/fluid-and-crystallized-intelligence 

General formula = y (dependent variable) ~ x (independent variable) + covariates
  Data fit 1 formula = thickness ~ eq + age + sex + handedness 
  Data fit 2 formula = thickness ~ eq + age + sex + handedness + fsiq 

```{r}
library(lme4)
library(sjPlot)
library(jtools)
library(devtools)
#library(cli)
#devtools::install_github("jacob-long/jtools")

#INSULA
lins_fit1 <- lm(lins_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
lins_fit2 <- lm(lins_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2) 
lins_fit1_table <- tab_model(lins_fit1)
lins_fit2_table <- tab_model(lins_fit2)
summ(lins_fit1) 
summ(lins_fit2)

rins_fit1 <- lm(rins_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
rins_fit2 <- lm(rins_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2)
rins_fit1_table <- tab_model(rins_fit1) 
rins_fit2_table <-tab_model(rins_fit2)
summ(rins_fit1) 
summ(rins_fit2)

#INFERIOR FRONTAL GYRUS
lifg_fit1 <- lm(lifg_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
lifg_fit2 <- lm(lifg_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2)
lifg_fit1_table <-tab_model(lifg_fit1) 
lifg_fit2_table <-tab_model(lifg_fit2)
summ(lifg_fit1) 
summ(lifg_fit2)

rifg_fit1 <- lm(rifg_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
rifg_fit2 <- lm(rifg_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2)
rifg_fit1_table <-tab_model(rifg_fit1) 
rifg_fit2_table <-tab_model(rifg_fit2)
summ(rifg_fit1) 
summ(rifg_fit2) 

#MEDIAL PREFRONTAL CORTEX REGIONS 
  #Rostral Anterior Cingulate Cortex
lracc_fit1 <- lm(lracc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
lracc_fit2 <- lm(lracc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2)
lracc_fit1_table <-tab_model(lracc_fit1) 
lracc_fit2_table <-tab_model(lracc_fit2) 
summ(lracc_fit1) 
summ(lracc_fit2)

rracc_fit1 <- lm(rracc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
rracc_fit2 <- lm(rracc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2)
rracc_fit1_table <-tab_model(rracc_fit1) 
rracc_fit2_table <-tab_model(rracc_fit2)
summ(rracc_fit1) 
summ(rracc_fit2)

  #ventral medial prefrontal cortex
lvmpfc_fit1 <- lm(lvmpfc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
lvmpfc_fit2 <- lm(lvmpfc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2)
lvmpfc_fit1_table <-tab_model(lvmpfc_fit1) 
lvmpfc_fit2_table <-tab_model(lvmpfc_fit2)
summ(lvmpfc_fit1) 
summ(lvmpfc_fit2)

rvmpfc_fit1 <- lm(rvmpfc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c, data = data2)
rvmpfc_fit2 <- lm(rvmpfc_etiv_mean_c ~ 1 + eq_mean_c + age_mean_c + sex_mean_c + handedness_mean_c + fsiq_mean_c, data = data2)
rvmpfc_fit1_table <-tab_model(rvmpfc_fit1) 
rvmpfc_fit2_table <-tab_model(rvmpfc_fit2) 
summ(rvmpfc_fit1) 
summ(rvmpfc_fit2) 
```
Notes from conversation with Marco: 
Checking for correlations between predictor variables. If VIF is <5, not of concern.Tells if model poorly fitting 
```{r}
summ(lins_fit1,vifs=TRUE,pvals=TRUE)
summ(lins_fit2,vifs=TRUE,pvals=TRUE)
summ(rins_fit1,vifs=TRUE,pvals=TRUE)
summ(rins_fit2,vifs=TRUE,pvals=TRUE)

summ(lifg_fit1,vifs=TRUE,pvals=TRUE)
summ(lifg_fit2,vifs=TRUE,pvals=TRUE)
summ(rifg_fit1,vifs=TRUE,pvals=TRUE)
summ(rifg_fit2,vifs=TRUE,pvals=TRUE)

summ(lracc_fit1,vifs=TRUE,pvals=TRUE)
summ(lracc_fit2,vifs=TRUE,pvals=TRUE)
summ(rracc_fit1,vifs=TRUE,pvals=TRUE)
summ(rracc_fit2,vifs=TRUE,pvals=TRUE)

summ(lvmpfc_fit1,vifs=TRUE,pvals=TRUE)
summ(lvmpfc_fit2,vifs=TRUE,pvals=TRUE)
summ(rvmpfc_fit1,vifs=TRUE,pvals=TRUE)
summ(rvmpfc_fit2,vifs=TRUE,pvals=TRUE)
```
Step 4. Plotting Results 
```{r}
library("ggplot2")

ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}
 
ggplotRegression(lifg_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Left Inferior Frontal Gyrus Cortical Thickness")

ggplotRegression(rifg_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Right Inferior Frontal Gyrus Cortical Thickness")

ggplotRegression(lins_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Left Insula Cortical Thickness")

ggplotRegression(rins_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Right Insula Cortical Thickness")

ggplotRegression(lracc_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Left Rostral Anterior Cingulate Cortex Cortical Thickness")

ggplotRegression(rracc_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Right Rostral Anterior Cingulate Cortex Cortical Thickness")

ggplotRegression(lvmpfc_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Left Ventral Medial Prefrontal Cortex Cortical Thickness")

ggplotRegression(rvmpfc_fit1) + 
  xlab("Empathizing Quotient") + 
  ylab("Right Ventral Medial Prefrontal Cortex Cortical Thickness")
```

Quick test to see directionality of age and gender effects on EQ 
```{r}
library(lme4)
library(sjPlot)
library(jtools)
library(devtools)

lins_fit3 <- lm(eq_mean_c ~ 1 + age_mean_c + lins_etiv_mean_c + sex_mean_c + handedness_mean_c, data = data2)

lins_fit4 <- lm(eq_mean_c ~ 1 + sex_mean_c + age_mean_c + lins_etiv_mean_c + handedness_mean_c, data = data2)

summ(lins_fit3) 
summ(lins_fit4)

library("ggplot2")

ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}

ggplotRegression(lins_fit3) + 
  xlab("Age") + 
  ylab("Empathizing Quotient")

ggplotRegression(lins_fit3) + 
  xlab("Gender") + 
  ylab("Empathizing Quotient")

```

